---
title: "Reproducible Research: Peer Assessment 1"
author: "Hiyong Byun"
date: "26/05/2019"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

## Loading and preprocessing the data
```{r}
library(readr)
library(dplyr)
activity <- read_csv("activity.csv")
activity <- na.omit(activity)
```

## What is mean total number of steps taken per day?
Histogram of the total number of steps taken each day
```{r}
totalstepsperday <- tapply(activity$steps, activity$date, sum)
hist(totalstepsperday, breaks = 30, xlab="Total Steps per Day", main="Histogram of Total Steps per Day")
```

Mean and median of the total number of steps taken per day
```{r}
mean(totalstepsperday)
median(totalstepsperday)
```

## What is the average daily activity pattern?
Time series plot of the average number of steps taken per day
```{r}
averagestepsinterval <- tapply(activity$steps, activity$interval, mean)
plot(averagestepsinterval, type="l", xlab="5 Minute Intervals", ylab="Average Number of Steps", main="Time Series of Average Number of Steps")
```

The 5-minute interval, on average across all the days in the dataset that contains the maximum number of steps
```{r}
averagestepsinterval[which.max(averagestepsinterval)]
```

## Imputing missing values
```{r}
activityimputed <- read_csv("activity.csv", col_types = cols(date = col_date(format = "%Y-%m-%d")))
```

Number of missing values
```{r}
sum(is.na(activityimputed$steps))
```

Calculate average number of steps for each interval
```{r}
intervalstepsmean <- tapply(activity$steps, activity$interval, mean)
```

Replace NA values with average number of steps for each interval
```{r}
intervalstepsmeandf <- as.data.frame(intervalstepsmean)
intervalstepsmeandf$interval <- rownames(intervalstepsmeandf)
for (i in 1:nrow(activityimputed)) {
        if (is.na(activityimputed$steps[i])) {
                x <- activityimputed$interval[i]
                activityimputed$steps[i] <- intervalstepsmeandf[intervalstepsmeandf$interval==x,1]
        }
}
```

Histogram of the total number of steps taken each day after missing values are imputed
```{r}
stepsimputed <- tapply(activityimputed$steps, activityimputed$date, sum)
hist(stepsimputed, breaks = 30, xlab="Total Steps per Day", main="Histogram of Total Steps per Day")
```

## Are there differences in activity patterns between weekdays and weekends?
Indicate weekday or weekend in new column in data frame
```{r}
for (i in 1:nrow(activityimputed)) {
        if (weekdays(as.POSIXlt(activityimputed$date[i])) == c("Saturday","Sunday")) {
                activityimputed$day[i] <- "weekend"
        } else {
                activityimputed$day[i] <- "weekday"
        }
}
```

Split data frame into weekday and weekend
```{r}
activityweekday <- filter(activityimputed, day=="weekday")
activityweekend <- filter(activityimputed, day=="weekend")
meanstepsweekday <- tapply(activityweekday$steps, activityweekday$interval, mean)
meanstepsweekend <- tapply(activityweekend$steps, activityweekend$interval, mean)
```

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
par(mfcol=c(2,1))
plot(meanstepsweekday, type="l", xlab="5 Minute Intervals", ylab="Average Steps Weekday")
plot(meanstepsweekend, type="l", xlab="5 Minute Intervals", ylab="Average Steps Weekend")
```
